---
title: "Creating Graphs in R with `ggplot2`"
author: "Ali El Hage Youssef"
date: "2023-07-23"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(here)
```

```{r}
ca_np <- read_csv(here("data","ca_np.csv"))
ci_np <- read_xlsx(here("data", "ci_np.xlsx"))
```

#Note: If you want to explicitly read in an .xlsx or .xls file, you can use read_xlsx() or read_xls() instead. read_excel() will #make its best guess about which type of Excel file you’re reading in, so is the generic form


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```
```{r}
gg_base <- ggplot(data = ci_np, aes( x = year, y = visitors))
gg_base +
  geom_point()
```
```{r}
gg_base +
  geom_col()
```

```{r}
gg_base +
  geom_area()
```

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
    
  )
```
```{r}
gg_base +
  geom_point( 
    color = "purple",
    pch = 17,
    size = 4,
    alpha = 0.5
    )
```

When we want to customize a graph element based on a variable’s characteristic or value, add the argument within aes() in the appropriate geom_*() layer


```{r}
gg_base +
  geom_point( 
  aes(
  size = visitors,
  color = visitors
),
alpha = 0.5
    )
```
In the example above, notice that the two arguments that do depend on variables are within aes(), but since alpha = 0.5 doesn’t depend on a variable then it is outside the aes() but still within the geom_point() layer.

```{r}
gg_base +
  geom_col( aes(fill = visitors))
```
## ggplot2 complete themes

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()

```
## Updating axis labels and titles

Use labs()

```{r}
gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual Park Visitors",
    title = " Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  )
```

```{r}
library(scales)
```
```{r}
gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual Park Visitors",
    title = " Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  ) +
  scale_y_continuous(labels = comma )
```


## Combining compatible geoms

As long as the geoms are compatible, we can layer them on top of one another to further customize a graph.

```{r}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5
             )
```

```{r}
gg_base +
  geom_col(fill = "orange",
           color = "green") +
  geom_line(color = "purple")
```
## Multi-series ggplot graphs







